---
output:
  pdf_document:
    keep_tex: true
    includes: 
      in_header: header.tex
bibliography: lib.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F, 
                      root.dir = "~/Dropbox/Schools/Public code/4 - Output")
knitr::opts_knit$set(root.dir = "~/Dropbox/Schools/Public code/4 - Output")
source("~/Dropbox/Schools/Public code/4 - Output/results.R")
```

\begin{center} 
    \textbf{\scshape \LARGE Supplemental Information}\\  \vspace{2mm}
    {\large Passing the Test}\\ \vspace{2mm} 
    {\large Alyssa Bilinski \footnote{Contact: abilinski@g.harvard.edu},\footnote{Data and code are available on GitHub (\href{https://github.com/abilinski/BackToSchool2}{link}).} $\cdot$ Joshua A. Salomon $\cdot$ Andrea Ciaranello \newline John Giardina $\cdot$ Meagan C. Fitzpatrick} \\
\end{center}
    

# Figures

```{r figs1, fig = T, fig.width = 8, fig.height = 7, fig.cap = "Sensitivity analysis - elementary schools by case type.  Average number of total secondary transmissions over 30 days (outside of the index case's household) following a single introduction into an elementary school community.  These include both transmission directly from the index case, as well as from secondary and tertiary cases.  The x-axes varies the level of prevention measure uptake, with low uptake assuming minimal interventions and high uptake assuming intensive interventions.  Line colors correspond to scheduling strategies."}

print(a[[3]])

```


```{r figs3, fig = T, fig.width = 8, fig.height = 10, fig.cap = "Sensitivity analyses (elementary schools) -- average number of total secondary transmissions over 30 days (outside of the index case's household) following a single introduction into a school community.  The columns vary the level of prevention measure uptake, with low uptake assuming minimal interventions and high uptake assuming intensive interventions.  Line colors correspond to scheduling strategies."}

ggarrange(a[[1]], esa[[1]],esc[[1]], esb[[1]], esd[[1]], ncol = 1, legend = "right")

```



```{r figs4, fig = T, fig.width = 8, fig.height = 10, fig.cap = "Sensitivity analyses (elementary schools) -- average number of total secondary transmissions over 30 days (outside of the index case's household) following a single introduction into a school community.  The columns vary the level of prevention measure uptake, with low uptake assuming minimal interventions and high uptake assuming intensive interventions.  Line colors correspond to scheduling strategies."}

ggarrange(b[[1]], hsa[[1]], hsc[[1]], hsb[[1]], ncol = 1, legend = "right")

```

\blandscape

```{r fig2, fig = T, fig.width = 8, fig.height = 7, fig.cap = "Average number of clinically symptomatic cases in staff and students over 30 days following a single introduction into a school community.  These include both transmission directly from the index case, as well as from secondary and tertiary cases.  The top panel shows elementary schools, where children are assumed to be less susceptible and less infectious, while the bottom panel shows high schools.  Note that axes differ across rows.  The x-axes varies the level of prevention measure uptake, with low uptake assuming minimal interventions and high uptake assuming intensive interventions.  Line colors correspond to scheduling strategies."}

ggarrange(a[[2]],b[[2]], common.legend = T, ncol = 1, legend = "right")

```

\elandscape

```{r figs2, fig = T, fig.width = 8, fig.height = 7, fig.cap = "Sensitivity analysis - high schools by case type.  Average number of total secondary transmissions over 30 days (outside of the index case's household) following a single introduction into an elementary school community.  These include both transmission directly from the index case, as well as from secondary and tertiary cases.  The x-axes varies the level of prevention measure uptake, with low uptake assuming minimal interventions and high uptake assuming intensive interventions.  Line colors correspond to scheduling strategies."}

print(b[[3]])

```

\blandscape

```{r, fig = T, fig.width = 11, fig.height = 8, fig.cap = "Cumulative incidence over 8 weeks in elementary schools across different levels of out-of-school mixing.  The x-axis shows the average daily community incidence per 100,000 population.  The y-axis shows cumulative incidence over 8 weeks.  Columns denote different isolation, quarantine, vaccination, and detection strategies, while rows show different population subgroups."}
 
print(a4[[3]])

```

\elandscape

\blandscape

```{r, fig = T, fig.width = 11, fig.height = 8, fig.cap = "Cumulative incidence over 8 weeks in high schools across different levels of out-of-school mixing.  The x-axis shows the average daily community incidence per 100,000 population.  The y-axis shows cumulative incidence over 8 weeks.  Columns denote different isolation, quarantine, vaccination, and detection strategies, while rows show different population subgroups."}
 
print(b4[[3]])

```

\elandscape


\clearpage

# Model

We use a Framework for Reconstructing Epidemiological Dynamics (FRED) to generate household structures [@wheaton_us_2014].  For computational simplicity, we used Maryland as a representative state, as sibling structure (the main parameter of interest) did not appear sensitive to location.

\begin{tabular}{cccc}
Maryland Elementary & Connecticut Elementary & Mississippi Elementary  & Texas Elementary  \\
\includegraphics[width=40mm]{/Users/abilinski/Dropbox/Schools/Public code/0 - Synthetic Populations/2 - Output/sibling_heatmap_Maryland.png} &
      \includegraphics[width=40mm]{/Users/abilinski/Dropbox/Schools/Public code/0 - Synthetic Populations/2 - Output/sibling_heatmap_Connecticut.png} &
      \includegraphics[width=40mm]{/Users/abilinski/Dropbox/Schools/Public code/0 - Synthetic Populations/2 - Output/sibling_heatmap_Mississippi.png} &
      \includegraphics[width=40mm]{/Users/abilinski/Dropbox/Schools/Public code/0 - Synthetic Populations/2 - Output/sibling_heatmap_Texas.png} 
    \end{tabular}

\begin{tabular}{cccc}
Maryland HS & Connecticut HS & Mississippi HS  & Texas HS  \\
\includegraphics[width=40mm]{/Users/abilinski/Dropbox/Schools/Public code/0 - Synthetic Populations/2 - Output/sibling_heatmap_Maryland_HS.png} &
      \includegraphics[width=40mm]{/Users/abilinski/Dropbox/Schools/Public code/0 - Synthetic Populations/2 - Output/sibling_heatmap_Connecticut_HS.png} &
      \includegraphics[width=40mm]{/Users/abilinski/Dropbox/Schools/Public code/0 - Synthetic Populations/2 - Output/sibling_heatmap_Mississippi_HS.png} &
      \includegraphics[width=40mm]{/Users/abilinski/Dropbox/Schools/Public code/0 - Synthetic Populations/2 - Output/sibling_heatmap_Texas_HS.png} 
    \end{tabular}

\clearpage



# References